Cartogram Mapping and its Application to Cancer Data Visualization

### 2019-09-24 review_article.Rmd
re-organised some material and caption for interactivity examples

Choropleths are uni-variate displays, they do not inherently support multi-dimensionality.
This makes it difficult to pair demographic, environmental and other factors with spatial distributions.
@MMST present linked micromap plots as a solution to the issue of multi-dimensionality to visually link geographic and statistical data.
This design is widely used in online atlas displays, it groups areas based on their value for one variable, and uses  additional columns to provide displays that contrast the areas in each group by other variables.
The display juxtaposes of choropleth maps and statistical plots;
it shows one map per group of the key separating variable, in a row with each additional statistical plot.
Linked micromaps predominantly utilize the choropleth map for displays of spatial distributions.
It is one of several alternative displays that allow maps to become bi-variate displays.

@@

@@

There are now many suitable methods and mediums to communicate health data.
The choice between them should always be driven by the intended message to be communicated and the needs of the audience.

@CPISACA provide three questions that are generally asked when using a disease map:

@@

```{r animating, fig.cap = "Animations in publicly available choropleth cancer maps. [a. Map of Cancer Mortality Rates in Spain](http://www.geeitema.org/AtlasET/atlas.jsp?causa=e27MEPOC)(Controls for indicator, sex, age groups, continents, and cancer types) , [b. Breast Cancer Mortality in Canada](http://www.ehatlas.ca/light-pollution/maps/breast-cancer-mortality) (Controls for indicator, age, year, and geographical areas)", fig.width = 12, fig.height = 9}
a1 <- png::readPNG("cancer_map_images/animation/23.animation.png")
a2 <- png::readPNG("cancer_map_images/animation/2.animation.png")

>>>>

Choropleths are uni-variate displays, they do not inherently support multi-dimensionality.
This makes it difficult to pair demographic, environmental and other factors with spatial distributions.

@@

@MMST present linked micromap plots as a solution to the issue of multi-dimensionality to visually link geographic and statistical data.
This design is widely used in online atlas displays, it groups areas based on their value for one variable, and uses  additional columns to provide displays that contrast the areas in each group by other variables.
The display juxtaposes of choropleth maps and statistical plots;
it shows one map per group of the key separating variable, in a row with each additional statistical plot.
Linked micromaps predominantly utilize the choropleth map for displays of spatial distributions.
It is one of several alternative displays that allow maps to become bivariate displays. (XXX Maybe also refer to Petra Kunert's visumap package since bivariate map displays has been raised here)

@@

<!-- this is vague and uninformative 
There are now many suitable methods and mediums to communicate health data.
The choice between them should always be driven by the intended message to be communicated and the needs of the audience.-->

Interactivity and animation add the ability to display more information in the same display, without overloading a single display. It is also a way to invite the viewer to engage with the display.

XXX Define interactivity, define animation

XXX summarise literature on interactivity with maps, and also animation.

Thinking about interactivity in terms of atlas purpose, it is useful to identify the primary questions that might be asked (or communicated). @CPISACA provide three questions that are generally asked when using a disease map:

@@

```{r animating, fig.cap = "Interactivity and animations in publicly available choropleth cancer maps. [a. Map of Cancer Mortality Rates in Spain](http://www.geeitema.org/AtlasET/atlas.jsp?causa=e27MEPOC)(Controls for indicator, sex, age groups, continents, and cancer types) , [b. Breast Cancer Mortality in Canada](http://www.ehatlas.ca/light-pollution/maps/breast-cancer-mortality) (Controls for indicator, age, year, and geographical areas)", fig.width = 12, fig.height = 9}
a1 <- png::readPNG("cancer_map_images/animation/23.animation.png")
a2 <- png::readPNG("cancer_map_images/animation/2.animation.png")
